Iterative Monomorphisation

### 2025-02-13 iterative_monomorphisation.tex
fix url
too big for my server
cites

B\"ohme's iterative approach is implemented as part of Isabelle/HOL's SMT (satisfiability modulo theories) integration~\cite[Chapter 2]{sb-phd}. This implementation is also used by Sledgehammer~\cite{judgement,hammer} to interface with superposition based automatic theorem provers. However, it is is documented only as a single subsection in B\"ohme's PhD thesis~\cite[Section 2.2.1]{sb-phd}.
A similar algorithm appears to be implemented in the \texttt{MESON} tactic~\cite{harrison-1996} of HOL Light, but it is undocumented.

@@

We implemented iterative monomorphisation in Zipperposition~\cite{zipp}, a higher order prover written in OCaml. Although Zipperposition is polymorphic, it uses the monomorphic prover E~\cite{e} as a backend. Thanks to our work, E can now be used with polymorphic problems. Moreover, our implementation can be used as a preprocessor to interface with other stand-alone provers.

@@

The monomorphisation algorithm is parametrised by many bounds. The first part of the evaluation process is to find appropriate values for these bounds. The second part compares the performance of Zipperposition with and without E on polymorphic problems. The third part compares the performance of different provers on polymorphic problems and their monomorphised counterparts.
Our raw evaluation data is available online.%
\footnote{\url{https://nekoka-project.github.io/pubs/monomorphization_data.zip}}

>>>>

B\"ohme's iterative approach is implemented as part of Isabelle/HOL's SMT (satisfiability modulo theories) integration~\cite[Chapter 2]{sb-phd}. This implementation is also used by Sledgehammer~\cite{judgement,paulson-blanchette-2010} to interface with superposition based automatic theorem provers. However, it is is documented only as a single subsection in B\"ohme's PhD thesis~\cite[Section 2.2.1]{sb-phd}.
A similar algorithm appears to be implemented in the \texttt{MESON} tactic~\cite{harrison-1996} of HOL Light, but it is undocumented.

@@

We implemented iterative monomorphisation in Zipperposition~\cite{vukmirovic-et-al-2021}, a higher order prover written in OCaml. Although Zipperposition is polymorphic, it uses the monomorphic prover E~\cite{e} as a backend. Thanks to our work, E can now be used with polymorphic problems. Moreover, our implementation can be used as a preprocessor to interface with other stand-alone provers.

@@

The monomorphisation algorithm is parametrised by many bounds. The first part of the evaluation process is to find appropriate values for these bounds. The second part compares the performance of Zipperposition with and without E on polymorphic problems. The third part compares the performance of different provers on polymorphic problems and their monomorphised counterparts.
Our raw evaluation data is available online.%
\footnote{\url{TODO}}
